```{r eval=FALSE}

# interactive 8 128G 10:00:00 genD
# conda activate spatial-2024

# note: don't update packages!!!!!!!

library(Voyager)
library(SpatialFeatureExperiment)
library(rjson)
library(Matrix)
library(SFEData)
library(RBioFormats)


library(SingleCellExperiment)
library(SpatialExperiment)
library(ggplot2)
library(stringr)
library(scater) 
library(scuttle)
library(BiocParallel)
library(BiocSingular)
library(bluster)
library(scran)
library(patchwork)
library(cowplot)
theme_set(theme_cowplot())
library(Seurat)
library(tidyverse)


```

```{r eval = FALSE}

setwd('~/collab/Valiente_2024/individual/')
data_dir <- 'data/'
fig_dir <- 'figures/'


# load one SFE dataset

cur_sample <- '134990_left'

sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))

# re-load the subset dataset
sfe_sub <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))
sfe_tumor <- sfe_sub[,sfe_sub$cell_group == 'Tumor']


```


Calculate signatures

```{r eval=FALSE}

#-----------------------------------------------------------------
# Broad cell type annotation based on marker genes + 
# transcriptomic clusters
#-----------------------------------------------------------------

ct_df <- data.frame(
    cluster = c(1:11),
    cell_group = c('Tumor', 'CNS', 'Immune', 'Vasculature', 'CNS', 'Immune', 'CNS', 'Vasculature', 'Vasculature', 'CNS', 'Immune')
)

tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group


```

Perform additional clustering to identify micromets 

Slightly different settings for each sample so 
copying this block to run it sample by sample 

sample 134990_left

```{r eval=FALSE}

set.seed(22)

#---------------------------------------------------------------
# step 1: KNN clustering based on spatial coordinates 
#---------------------------------------------------------------

sfe2 <- imcRtools::buildSpatialGraph(
    sfe, img_id = "SampleID",
    type = 'knn', 
    coords = c('x_centroid', 'y_centroid'),
    k=10,
)

# aggregate neighbors by expression levels 
sfe2 <- imcRtools::aggregateNeighbors(
    sfe2,
    colPairName="knn_interaction_graph",
    aggregate_by = "expression",
    assay_type='logcounts'
)

#---------------------------------------------------------------
# step 2: K-means clustering based on neighbor aggregates
#---------------------------------------------------------------

# k values to sweep
k_list <- 2:6

# kmeans clustering of the aggregated expression matrix
for(cur_k in k_list){
    cur_cluster <- kmeans(sfe2$mean_aggregatedExpression, centers = cur_k)
    cluster_name <- paste0('niche_neighbors_knn', cur_k)
    colData(sfe)[,cluster_name] <- factor(cur_cluster$cluster)
}

# plot the results for the different K values
plot_list <- list()
for(cur_k in k_list){
    cluster_name <- paste0('niche_neighbors_knn', cur_k)
    plot_list[[cluster_name]] <- plotSpatialFeature(
        sfe, cluster_name, 
        colGeometryName = "cellSeg") + Seurat::NoLegend() + 
        ggtitle(paste0('k = ', cur_k))
}

pdf(paste0(cur_sample, '/figures/', 'spatial_niches.pdf'), width=12, height=6)
wrap_plots(plot_list, ncol=5)
dev.off()

cur_k <- 2
cluster_name <- paste0('niche_neighbors_knn', cur_k)
p <- plotSpatialFeature(
    sfe, cluster_name, 
    colGeometryName = "cellSeg") + 
    ggtitle(paste0('k = ', cur_k))

pdf(paste0(cur_sample, '/figures/', 'spatial_niches_knn', cur_k,'.pdf'), width=6, height=6)
p
dev.off()

#------------------------------------------------------------
# step3: subset the micromet cluster, then perform 
# leiden clustering to identify individual micromets
#
# continue here for the cur sample, 
#------------------------------------------------------------
table(colData(sfe)$niche_neighbors_knn2)


# subset the CNS cells 
sfe_sub <- sfe[,sfe$cell_group != "CNS"]

# subset the K-means cluster containing the micromets
sfe_sub <- sfe_sub[,colData(sfe_sub)$niche_neighbors_knn2 == 1]

# build a new KNN graph
sfe_sub2 <- imcRtools::buildSpatialGraph(
    sfe_sub, img_id = "SampleID",
    type = 'knn', 
    coords = c('x_centroid', 'y_centroid'),
    k=3,
)

# perform leiden community detection
sfe_sub2 <- imcRtools::detectCommunity(
    sfe_sub2, 
    colPairName = "knn_interaction_graph",
    BPPARAM = list(resolution = 1)
)
colData(sfe_sub)$spatial_community <- colData(sfe_sub2)$spatial_community

# compute the cluster centroids (for plotting labels)
centroid_df <- colData(sfe_sub) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))

# plot the clusters in space
p <- plotSpatialFeature(
        sfe_sub, 'spatial_community', 
        colGeometryName = "cellSeg") +
        ggtitle("Clustering Micromets") + scale_fill_discrete() + 
        geom_text_repel(
            data = centroid_df,
            mapping = aes(x=x, y=y),
            label = centroid_df$spatial_community,
            max.overlaps=Inf,
            size=3
        )


pdf(paste0(cur_sample, '/figures/', 'spatial_community.pdf'), width=8, height=8)
p
dev.off()

#-----------------------------------------------------------------
# Step4: Refine the clustering by merging, filtering, refining
#-----------------------------------------------------------------
#colData(sfe_sub)$spatial_community<- sfe_sub$spatial_community_orig

colData(sfe_sub)$spatial_community_orig <- sfe_sub$spatial_community

# clusters to merge
merge_clusters <- list(
    '1' = c(1, 29, 24, 19, 44, 5, 6, 7, 30, 31, 45, 42, 2, 3, 41, 43, 21, 22, 20, 18, 17, 54, 26, 25, 28, 27, 24, 23),
    '8' = c(51, 8, 46),
    '36' = c(36, 37)
)

for(x in names(merge_clusters)){
    print(x)
    cur_merge <- merge_clusters[[x]]
    cur_clusters <- as.numeric(colData(sfe_sub)$spatial_community)
    updated_clusters <- ifelse(
        cur_clusters %in% cur_merge, as.numeric(x), cur_clusters 
    ) 
    colData(sfe_sub)$spatial_community <- updated_clusters
}
table(sfe_sub$spatial_community)

# re-number the clusters based on how big they are 
tmp <- table(sfe_sub$spatial_community)
tmp2 <- names(tmp)

tmp_order <- order(-tmp); 

tmp2 <- names(tmp[tmp_order])
names(tmp2) <- as.character(1:length(tmp2))


ix <- match(sfe_sub$spatial_community, as.numeric(tmp2))
sfe_sub$spatial_community <- names(tmp2)[ix]

sfe_sub$spatial_community <- factor(
    as.numeric(sfe_sub$spatial_community),
    levels = 1:length(tmp)
)

# compute the cluster centroids (for plotting labels)
centroid_df <- colData(sfe_sub) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))

# plot the clusters in space
p <- plotSpatialFeature(
        sfe_sub, 'spatial_community', 
        colGeometryName = "cellSeg") +
        ggtitle("Clustering Micromets") + scale_fill_discrete() + 
        geom_text(
            data = centroid_df,
            mapping = aes(x=x, y=y),
            label = centroid_df$spatial_community,
            size=3
        )


pdf(paste0(cur_sample, '/figures/', 'spatial_community_merged.pdf'), width=8, height=8)
p
dev.off()



#-----------------------------------------------------------------
# Plot cell-type compositions of the micromets:
#-----------------------------------------------------------------

# get the color scheme:
p <- plotSpatialFeature(sfe_sub, "cell_group", colGeometryName = "cellSeg") 

pdf(paste0(cur_sample, '/figures/', 'leiden_spatial_merged_celltypes.pdf'), width=8, height=8)
p
dev.off()

# get the color scheme
g <- ggplot_build(p)
color_df <- g$data[[1]] %>% dplyr::select(c('group', 'fill')) %>% dplyr::distinct()
cp <- color_df$fill 
names(cp) <- unique(sfe_sub$cell_group)



plot_df <- table(sfe_sub$spatial_community, sfe_sub$cell_group) %>% 
    as.data.frame() %>%
    reshape2::melt()
colnames(plot_df)[1:2] <- c('spatial_community', 'cell_group')


p <- plot_df %>% 
    ggplot(aes(x=spatial_community, y=value, fill=cell_group)) +
    geom_bar(position='fill', stat='identity') +
    ylab('Proportion') + 
    xlab('Met group') +
    scale_fill_manual(values=cp)


pdf(paste0(cur_sample, '/figures/', 'spatial_community_proportions.pdf'), width=10, height=4)
p
dev.off()

# make another plot but order by tumor fraction
tmp <- plot_df %>% 
    group_by(spatial_community) %>% 
    summarise(total = sum(value))

tmp2 <- plot_df %>% subset(cell_group == 'Tumor')
frac <- tmp2$value / tmp$total
names(frac) <- as.character(tmp$spatial_community)

plot_df$spatial_community <- factor(
    as.character(plot_df$spatial_community),
    levels = as.character(names(frac)[order(-frac)])
)

p <- plot_df %>% 
    ggplot(aes(x=spatial_community, y=value, fill=cell_group)) +
    geom_bar(position='fill', stat='identity') +
    ylab('Proportion') + 
    xlab('Met group') +
    scale_fill_manual(values=cp)


pdf(paste0(cur_sample, '/figures/', 'spatial_community_proportions_ordered.pdf'), width=10, height=4)
p
dev.off()



#-----------------------------------------------------------------
# Calculate Signatures
#-----------------------------------------------------------------


genes_signature = list(
    'micromets'= c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A'),
    'endothelia'= c('STAT3', 'FLT1', 'NOTCH1', 'NRP1', 'THSD4', 'PECAM1', 'CAV1', 'VWF'),
    'dormancy'= c('NDRG1', 'DYRK1A', 'ZEB2', 'THBS1', 'BMPR2', 'TGFBR3', 'AXL', 'BHLHE41', 'RHEB', 'NR2F1'),
    'stress_general'= c('HIF1A', 'H2AZ1', 'NFE2L2', 'HSP90AA1', 'ATF6', 'ATM', 'ATR', 'NRF2', 'ATF3'),
    'proliferation'= c('CDK4', 'PCNA', 'MKI67', 'TOP2A', 'CCNB2'),
    'stress_ATF6' = c('ATF6',	'HSP90AA1',	'XBP1',	'CIP2A'),
    'stress_DNA_damage' = c('H2AZ1',	'ATM',	'ATR'),
    'mTOR' = c('MTOR',	'RPTOR',	'RHEB',	'IGF1',	'TCF4',	'DNAAF1',	'SRPX',	'TFPI',	'DES',	'CD86'),
    'dormancy_expanded' = c('RHEB',	'NR2F1',	'SOX9',	'AXL',	'ATF6',	'BMPR2',	'TGFBR3',	'RPS6KA5',	'NDRG1',	'BHLHE41',	'MTOR',	'DYRK1A',	'ZEB2',	'MAPK14',	'THBS1'),
    'autophagy' = c('SQSTM1',	'BECN1',	'MAP1LC3B',	'CAV1',	'RPTOR',	'LAMP3'),
    'TGFb' = c('BMPR2',	'TGFBR3',	'TGFB1',	'TGFB2',	'TGFBI',	'SNAI1'),
    'cancer_cells' = c('EPCAM',	'KRT20',	'KRT7'),
    'chromosomal_instability' = c('PCNA',	'UBE2C',	'CCNB2',	'TOP2A')
)

# adjust the signatures to only include genes in the dataset 
for(sig in names(genes_signature)){
    cur_genes <- genes_signature[[sig]]
    genes_signature[[sig]] <- cur_genes[cur_genes %in% rowData(sfe)$Symbol]
}

meta <- colData(sfe_sub)

test <- UCell::ScoreSignatures_UCell(
    as(counts(sfe_sub), 'dgCMatrix'), 
    features=genes_signature
)
colnames(test) <- names(genes_signature)
colData(sfe_sub) <- cbind(colData(sfe_sub), test)

# save the dataset
saveRDS(sfe_sub, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))


```

# Analyze signatures

```{r eval=FALSE}

#-----------------------------------------------------------------
# Calculate Signatures (only in Tumor cells!!!)
#-----------------------------------------------------------------

genes_signature = list(
    'micromets'= c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A'),
    'dormancy'= c('NDRG1', 'DYRK1A', 'ZEB2', 'THBS1', 'BMPR2', 'TGFBR3', 'AXL', 'BHLHE41', 'RHEB', 'NR2F1'),
    'stress_general'= c('HIF1A', 'H2AZ1', 'NFE2L2', 'HSP90AA1', 'ATF6', 'ATM', 'ATR', 'NRF2', 'ATF3'),
    'proliferation'= c('CDK4', 'PCNA', 'MKI67', 'TOP2A', 'CCNB2'),
    'stress_ATF6' = c('ATF6',	'HSP90AA1',	'XBP1',	'CIP2A'),
    'stress_DNA_damage' = c('H2AZ1',	'ATM',	'ATR'),
    'dormancy_expanded' = c('RHEB',	'NR2F1',	'SOX9',	'AXL',	'ATF6',	'BMPR2',	'TGFBR3',	'RPS6KA5',	'NDRG1',	'BHLHE41',	'MTOR',	'DYRK1A',	'ZEB2',	'MAPK14',	'THBS1'),
    'autophagy' = c('SQSTM1',	'BECN1',	'MAP1LC3B',	'CAV1',	'RPTOR',	'LAMP3'),
    'chromosomal_instability' = c('PCNA',	'UBE2C',	'CCNB2',	'TOP2A')
)

# adjust the signatures to only include genes in the dataset 
for(sig in names(genes_signature)){
    cur_genes <- genes_signature[[sig]]
    genes_signature[[sig]] <- cur_genes[cur_genes %in% rowData(sfe)$Symbol]
}

sfe_tumor <- sfe_sub[,sfe_sub$cell_group == 'Tumor']
meta <- colData(sfe_tumor)

test <- UCell::ScoreSignatures_UCell(
    as(counts(sfe_tumor), 'dgCMatrix'), 
    features=genes_signature
)
colnames(test) <- names(genes_signature)
colData(sfe_tumor) <- cbind(colData(sfe_tumor), test)




# plot vln plots with seurat
X <- counts(sfe_tumor)
seurat_obj <- CreateSeuratObject(
    counts=X, meta.data=as.data.frame(colData(sfe_tumor))
)

# loop through each signature:
seurat_obj$plot_group <- seurat_obj$spatial_community

# remove very small groups
small_groups <- names(which(table(seurat_obj$plot_group) < 5))
seurat_obj <- seurat_obj[,! seurat_obj$plot_group %in% small_groups]

plot_list <- list()
for(cur_sig in names(genes_signature)){
    print(cur_sig)
    cur_df <- seurat_obj@meta.data[,c('spatial_community', cur_sig)]
    names(cur_df)[2] <- 'value'
    cur_df <- cur_df %>% group_by(spatial_community) %>%
        summarise(mean = mean(value)) %>%
        arrange(desc(mean))

    group_order <- as.character(cur_df$spatial_community)
    seurat_obj$plot_group <- factor(
        as.character(seurat_obj$plot_group),
        levels = group_order
    )

    plot_list[[cur_sig]] <- VlnPlot(
        seurat_obj, 
        group.by = 'plot_group',
        features = cur_sig,
        pt.size=0
    ) + NoLegend() + xlab('')

}

pdf(paste0(cur_sample, '/figures/', 'signature_vlnplot.pdf'), width=16, height=12)
wrap_plots(plot_list, ncol=2)
dev.off()


# micromet genes
plot_list <- list()
for(gene in genes_signature$micromet){

    plot_list[[gene]] <- plotSpatialFeature(
        sfe_sub, gene, colGeometryName = "cellSeg", dark=TRUE) 
}


pdf(paste0(cur_sample, '/figures/', 'signature_micromet_featureplots.pdf'), width=15, height=5)
wrap_plots(plot_list, ncol=5)
dev.off()


median(table(sfe_sub$spatial_community))
quantile(table(sfe_sub$spatial_community))

#--------------------------------------------------------------------#
# micromet-level signature plots (raincloud plot)
#--------------------------------------------------------------------#

library(gghalves)
library(ggbeeswarm) 
library(ggdist)
library(ggpp)

plot_df <- data.frame()
for(cur_sig in names(genes_signature)){

    cur_df <- as.data.frame(colData(sfe_tumor)[,c('spatial_community', cur_sig)])
    colnames(cur_df)[2] <- 'score'
    cur_df$signature <- cur_sig 
    plot_df <- rbind(plot_df, cur_df)
}

# compute averages
plot_df <- plot_df %>% 
    group_by(spatial_community, signature) %>%
    summarise(value = mean(score))

sig_orders <- plot_df %>%
    group_by(signature) %>%
    summarise(x = mean(value)) %>% 
    arrange(x) %>% .$signature

plot_df$signature <- factor(as.character(plot_df$signature), levels=sig_orders)

# add colors:
#plot_df$color <- as.character(cp[as.character(plot_df$spatial_community)])

# data frame for text boxes:
top_df <- plot_df %>% 
    group_by(signature) %>% 
    slice_max(n=5, order_by=value) %>%
    summarise(text=paste(spatial_community, collapse=','))

bot_df <- plot_df %>%
    group_by(signature) %>% 
    slice_min(n=5, order_by=value) %>%
    summarise(text=paste(spatial_community, collapse=','))

text_df <- rbind(top_df, bot_df)




p <- plot_df %>% 
    ggplot(aes(y=signature, x=value, fill=signature)) +
    stat_slab(side = "top", scale = 0.45,show.legend = F,
            position = position_nudge(y = 0.1),
            alpha=0.75,
            aes(fill_ramp = stat(level)),.width = c(.50, .95,1)) +
    geom_boxplot(
        width = .05,outlier.alpha=0, alpha=0.75,
        position = position_nudge(y = 0.05),show.legend = FALSE) +
    stat_dots(
        side = "bottom", scale = 0.45,show.legend = T,
        position = position_dodge(width = .8), color=NA
        ) +
    scale_fill_ramp_discrete(from='black', aesthetics = "fill_ramp") +
    guides(color="none", fill='none') +
    RotatedAxis() + xlab('') + 
    xlab('Signature expression score') +
    ylab('') +
    geom_label(
        data=top_df, inherit.aes=FALSE,
        aes(y=signature, x=1, label=text), size=3) + 
    geom_label(
        data=bot_df, inherit.aes=FALSE,
        aes(y=signature, x=0.3, label=text), size=3) + 
    theme(
        panel.border=element_rect(size=1, color='black', fill=NA),
        panel.grid.major.x = element_line(color='lightgrey', linewidth=0.25),
        axis.line.x = element_blank(),
        axis.line.y = element_blank()
    ) + NoLegend() + 
    scale_x_continuous(breaks = c(0.2, 0.4, 0.6, 0.8, 1.0), limits=c(0.2,1.1))
  #  ggforestplot::geom_stripes(aes(y=signature), inherit.aes=FALSE, data=plot_df) +


pdf(paste0(cur_sample, '/figures/', 'signature_distributions.pdf'), width=10, height=5)
p
dev.off()

```  

Compare signatures to each other (correlation)
at the micromet level (?)

```{r eval=FALSE}

signatures <- names(genes_signature)
sig1 <- 'micromets'

plot_list <- list()
cor_df <- data.frame()
for(sig1 in signatures){

    for(sig2 in signatures){

        cur_cor <- cor(cur_df$sig1, cur_df$sig2)
        cor_df <- rbind(cor_df, data.frame(cor=cur_cor, sig1=sig1, sig2=sig2))

        if(sig1 == sig2){next}
        if(paste0(sig2, '_', sig1) %in% names(plot_list)){
            plot_list[[paste0(sig1, '_', sig2)]] <- ggplot() + theme_void()
            next
        }

        print(paste0(sig1, sig2))
        cur_df <- subset(plot_df, signature %in% c(sig1, sig2))

        cur_df <- cur_df %>% 
            tidyr::pivot_wider(names_from=signature, values_from=value)
        cur_df <- cur_df[,c('spatial_community', sig1, sig2)]
        colnames(cur_df) <- c('spatial_community', 'sig1', 'sig2')

        p <- cur_df %>% 
            ggplot(aes(x=sig1, y=sig2, color=spatial_community)) + 
            geom_point() + 
            geom_smooth(
                inherit.aes=FALSE,
                aes(x=sig1, y=sig2),
                method='lm'
            ) + 
            ggpubr::stat_cor() + 
            xlab(sig1) + ylab(sig2) + theme(
                axis.line.y = element_blank(),
                axis.line.x = element_blank(),
                panel.border = element_rect(linewidth=0.5, fill=NA, color='black')
            )

        plot_list[[paste0(sig1, '_', sig2)]] <- p
    }
}


pdf(paste0(cur_sample, '/figures/', 'signatures_compare.pdf'), width=20, height=20)
wrap_plots(plot_list, ncol=length(signatures)) + plot_layout(guides='collect')
dev.off()


# Heatmap of correlations

cor_df$sig1 <- factor(as.character(cor_df$sig1), levels=signatures)
cor_df$sig2 <- factor(as.character(cor_df$sig2), levels=signatures)

p <- cor_df %>% 
    ggplot(aes(x=sig1, y=sig2, fill=cor)) + 
    geom_tile() + 
    coord_fixed() + RotatedAxis() + 
    scale_fill_gradient2(high='red', mid='white', low='blue', limits=c(-1,1)) + 
    geom_text(
        data = subset(cor_df, abs(cor) > 0.5), 
        aes(label=round(cor, 2))
    ) + xlab('') + ylab('')


pdf(paste0(cur_sample, '/figures/', 'signatures_compare_heatmap.pdf'), width=6, height=6)
p
dev.off()



```



Can I plot the micromets on the full dataset ? 

```{r eval=FALSE}

ix <- as.character(colData(sfe_sub)$spatial_community)
names(ix) <- rownames(colData(sfe_sub))
tmp <- ix[colnames(sfe)]
colData(sfe)$spatial_community <- ifelse(is.na(tmp), 'other', tmp)
colData(sfe)$spatial_community <- factor(
    as.character(colData(sfe)$spatial_community), 
    levels = c(levels(colData(sfe_sub)$spatial_community), 'other')
)

library(scales)
library(RColorBrewer)


cp <- hue_pal()(length(levels(sfe$spatial_community))-1) 
cp <- c(cp, 'lightgrey')
names(cp) <- levels(sfe$spatial_community)

# compute the cluster centroids (for plotting labels)
centroid_df <- colData(sfe_sub) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))

# plot the clusters in space
p <- plotSpatialFeature(
        sfe, 'spatial_community', 
        colGeometryName = "cellSeg") +
        ggtitle("Clustering Micromets") + 
        scale_fill_manual(values=cp) +
        geom_text(
            data = centroid_df,
            mapping = aes(x=x, y=y),
            label = centroid_df$spatial_community,
            size=2
        ) + NoLegend()


pdf(paste0(cur_sample, '/figures/', 'spatial_community_refined_bg.pdf'), width=10, height=8)
p
dev.off()


table(sfe_sub$spatial_community)



#-------------------------------------------------------#
# Zoom plot per micromet
#-------------------------------------------------------#



plot_groups <- levels(sfe_sub$spatial_community)

# define color palettes
cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
names(cp) <- levels(sfe_sub$plot_group)

# bounding box offset
offset <- 25

bbox_df <- data.frame()
plot_list <- list()
for(cur_group in plot_groups){

    print(cur_group)

    # define the bbox for this group
    cur_df <- colData(sfe_sub) %>% as.data.frame() %>%
        subset(spatial_community == cur_group)

    # current plotting group
    colData(sfe)$plot_group <- ifelse(
        colData(sfe)$spatial_community == cur_group & colData(sfe)$spatial_community != 'other',
        colData(sfe)$cell_group, 'other'
    )
    colData(sfe)$plot_group <- factor(
        as.character(colData(sfe)$plot_group),
        levels = c('Tumor', 'Vasculature', 'Immune', 'other')
    )

    # define bounding box
    bbox <- c(
        xmin = min(cur_df$x_centroid) - offset,
        ymin = min(cur_df$y_centroid) - offset,
        xmax = max(cur_df$x_centroid) + offset,
        ymax = max(cur_df$y_centroid) + offset
    )

    bbox_df <- rbind(bbox_df, bbox)

    # plot the clusters in space
    p <- plotSpatialFeature(
            sfe, 'plot_group', 
            colGeometryName = "cellSeg",
            bbox = bbox
        ) +
        ggtitle(paste0("Micromet ", cur_group)) + 
        labs(fill="") +
        scale_fill_manual(values=cp) + theme(
            plot.title = element_text(face='bold', hjust=0.5)
        ) + theme(
            panel.border=element_rect(fill=NA, linewidth=0.5, color='black')
        ) #+ ggspatial::annotation_scale()
        
    plot_list[[cur_group]] <- p

}
names(bbox_df) <- c('xmin', 'ymin', 'xmax', 'ymax')

pdf(paste0(cur_sample, '/figures/', 'spatial_community_micromets_zoom.pdf'), width=15, height=15)
wrap_plots(plot_list, ncol=10) + plot_layout(guides='collect')
dev.off()


# plot the full spatial section
cp <- hue_pal()(length(levels(sfe$spatial_community))-1) 
cp <- c(cp, 'lightgrey')
names(cp) <- levels(sfe$spatial_community)


# compute the cluster centroids (for plotting labels)
centroid_df <- colData(sfe_sub) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))

# scale bar length 
plot_len <- max(sfe$x_centroid) - min(sfe$x_centroid)
plot_len / 10

p_full <-  plotSpatialFeature(
        sfe, 'spatial_community', 
        colGeometryName = "cellSeg") +
        ggtitle("Clustering Micromets"
    ) + 
    geom_rect(
            data = bbox_df, inherit.aes=FALSE,
            aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),
            fill=NA, color='black', linewidth=0.25
    ) + 
    geom_text(
        data = centroid_df,
        mapping = aes(x=x, y=y),
        label = centroid_df$spatial_community,
        size=2
    ) +
    scale_fill_manual(values=cp) + NoLegend() + theme(
        plot.title= element_text(hjust=0.5, face='bold')
    ) # + ggspatial::annotation_scale()



# plot the full spatial region with all the boxes 
pdf(paste0(cur_sample, '/figures/', 'spatial_community_refined_bg_boxes.pdf'), width=8, height=8)
p_full
dev.off()


#-------------------------------------------------------#
# Zoom plots of the individual micromet genes
#-------------------------------------------------------#

gene_list <- genes_signature$micromets

plot_groups <- levels(sfe_sub$spatial_community)

# define color palettes
cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
names(cp) <- levels(sfe_sub$plot_group)

# only show expression in tumor cells:
plot_sfe <- sfe[,sfe$cell_group == 'Tumor']


# bounding box offset
offset <- 25

bbox_df <- data.frame()
plot_list <- list()
cluster_plot_list <- list()
for(cur_group in plot_groups){

    print(cur_group)
    cur_plot_list <- list()

    # define the bbox for this group
    cur_df <- colData(sfe_sub) %>% as.data.frame() %>%
        subset(spatial_community == cur_group)

    # current plotting group
    colData(sfe)$plot_group <- ifelse(
        colData(sfe)$spatial_community == cur_group & colData(sfe)$spatial_community != 'other',
        colData(sfe)$cell_group, 'other'
    )
    colData(sfe)$plot_group <- factor(
        as.character(colData(sfe)$plot_group),
        levels = c('Tumor', 'Vasculature', 'Immune', 'other')
    )

    # define bounding box
    bbox <- c(
        xmin = min(cur_df$x_centroid) - offset,
        ymin = min(cur_df$y_centroid) - offset,
        xmax = max(cur_df$x_centroid) + offset,
        ymax = max(cur_df$y_centroid) + offset
    )

    bbox_df <- rbind(bbox_df, bbox)

    p <- plotSpatialFeature(
        sfe, 'plot_group', 
        colGeometryName = "cellSeg",
        bbox = bbox
    ) +
    ggtitle(paste0("Micromet ", cur_group)) + 
    labs(fill="") +
    scale_fill_manual(values=cp) + theme(
        plot.title = element_text(face='bold', hjust=0.5),
        panel.border=element_rect(fill=NA, linewidth=0.5, color='black')
    ) + theme(
        legend.position='bottom'
    )
    cur_plot_list[[cur_group]] <- p

  
    for(cur_gene in gene_list){

        # plot the clusters in space
        p <- plotSpatialFeature(
                plot_sfe, cur_gene, 
                colGeometryName = "cellSeg",
                bbox = bbox
            ) + 
            ggtitle(cur_gene) + 
            labs(fill="") + theme(
                panel.border = element_rect(linewidth=0.5, color='black', fill=NA),
                plot.title = element_text(hjust=0.5, face='bold'),
                legend.position='bottom'
            )
            
        cur_plot_list[[cur_gene]] <- p
    }
    plot_list[[cur_group]] <- cur_plot_list

}
names(bbox_df) <- c('xmin', 'ymin', 'xmax', 'ymax')

pdf(paste0(cur_sample, '/figures/', 'signature_micromets_zoom.pdf'), width=15, height=5)

    for(cur_plots in plot_list){
        patch <- wrap_plots(cur_plots, ncol=6)
        print(patch)
    }

dev.off()







```


Modeling gene expression as a function of distance to vasculature

```{r eval=FALSE}


meta <- as.data.frame(colData(sfe_sub))
meta$bc <- rownames(meta)
spatial_groups <- levels(sfe_sub$spatial_community)

# test for micromet 
cur_group <- 15

distance_df <- data.frame()
for(cur_group in spatial_groups){

    print(cur_group)

    cur_meta <- subset(meta, spatial_community == cur_group)

    # calculate distances
    cur_coords <- cur_meta %>% dplyr::select(x_centroid, y_centroid) %>%
        as.matrix()
    cur_distances <- dist(cur_coords, method='euclidean', diag=TRUE) %>%
        as.matrix() %>%
        reshape2::melt() %>%
        as.data.frame() %>%
        dplyr::rename(dist=value) %>%
        subset(Var1 != Var2)

    # transfer cell type labels to the distance table
    tmp1 <- left_join(cur_distances,meta, by = c('Var1' = 'bc'))
    tmp2 <- left_join(cur_distances,meta, by = c('Var2' = 'bc'))
    cur_distances$group1 <- tmp1$cell_group
    cur_distances$group2 <- tmp2$cell_group

    cur_distances$group <- cur_group
    distance_df <- rbind(distance_df, cur_distances)
}

# look at expression of one gene
gene_list <- rownames(sfe)
cur_gene <- gene_list[1]
X <- counts(sfe_sub)
#X <- logcounts(sfe_sub)

model_list <- list()
summary_list <- list()
stat_df <- data.frame()
p_list <- c()
for(cur_gene in gene_list){
    print(cur_gene)

    # linear model to test the relationship between vascular distance and expression level
    model_df <- distance_df %>%
        subset(group1 == 'Tumor' & group2 == 'Vasculature') %>%
        group_by(Var1) %>%
        summarise(avg_dist = mean(dist))

    # add expression level 
    model_df$expression <- X[cur_gene, model_df$Var1]

    # add the cluster info 
    model_df$group <- colData(sfe_sub)[model_df$Var1, 'spatial_community']
    model_df$cluster <- colData(sfe_sub)[model_df$Var1, 'spatial_community']

    # set up vars
    y <- model_df$expression 
    x <- model_df$avg_dist 

    # linear model  (not using anymore)
    #cur_model_lm <- lm(y ~ x)
    # cur_model_lm <- lm(expression ~ avg_dist + group, data=model_df)
    # cur_summary <- summary(cur_model_lm)
    # p_list <- c(p_list, as.numeric(cur_summary$coefficients[,4][2]))

    # random effect model:
    res <- try(
        cur_model <- nlme::lme(fixed= y ~ x, random = ~ 1 | cluster, data = model_df)
    )
    if(inherits(res, "try-error")){next}
    cur_summary <- summary(cur_model)

    # pearson corr 
    cor(x,y)

    coefficients_table <- cur_summary$tTable
    coefficient_x <- coefficients_table["x", "Value"]
    p_value_x <- coefficients_table["x", "p-value"]

    cur_df <- data.frame(
        gene = cur_gene,
        pval = p_value_x,
        coef = coefficient_x,
        cor = cor(x,y)
    )
    stat_df <- rbind(stat_df, cur_df)

    model_list[[cur_gene]] <- cur_model
    summary_list[[cur_gene]] <- cur_summary
 
}

# calculate FDR-corrected p-vals 
stat_df$fdr <- p.adjust(stat_df$pval, 'fdr')

subset(stat_df, fdr < 0.05) %>% nrow # 20 significant hits

stat_df <- stat_df %>% mutate(log_fdr=-log(pval)) %>% arrange(desc(log_fdr))

write.table(stat_df, file=paste0(cur_sample, '/data/', cur_sample, '_tumor_vasc_distance_genes.tsv'), sep='\t', row.names=FALSE, quote=FALSE)



#-------------------------------------------------------#
# Zoom plots of the top 5 genes
#-------------------------------------------------------#

gene_list <- stat_df %>% 
    subset(cor > 0) %>% 
    slice_max(order_by=log_fdr, n=5) %>% .$gene

plot_groups <- levels(sfe_sub$spatial_community)

# define color palettes
cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
names(cp) <- levels(sfe_sub$plot_group)

# only show expression in tumor cells:
plot_sfe <- sfe[,sfe$cell_group == 'Tumor']


# bounding box offset
offset <- 25

bbox_df <- data.frame()
plot_list <- list()
cluster_plot_list <- list()
for(cur_group in plot_groups){

    print(cur_group)
    cur_plot_list <- list()

    # define the bbox for this group
    cur_df <- colData(sfe_sub) %>% as.data.frame() %>%
        subset(spatial_community == cur_group)

    # current plotting group
    colData(sfe)$plot_group <- ifelse(
        colData(sfe)$spatial_community == cur_group & colData(sfe)$spatial_community != 'other',
        colData(sfe)$cell_group, 'other'
    )
    colData(sfe)$plot_group <- factor(
        as.character(colData(sfe)$plot_group),
        levels = c('Tumor', 'Vasculature', 'Immune', 'other')
    )

    # define bounding box
    bbox <- c(
        xmin = min(cur_df$x_centroid) - offset,
        ymin = min(cur_df$y_centroid) - offset,
        xmax = max(cur_df$x_centroid) + offset,
        ymax = max(cur_df$y_centroid) + offset
    )

    bbox_df <- rbind(bbox_df, bbox)

    p <- plotSpatialFeature(
        sfe, 'plot_group', 
        colGeometryName = "cellSeg",
        bbox = bbox
    ) +
    ggtitle(paste0("Micromet ", cur_group)) + 
    labs(fill="") +
    scale_fill_manual(values=cp) + theme(
        plot.title = element_text(face='bold', hjust=0.5),
        panel.border=element_rect(fill=NA, linewidth=0.5, color='black')
    ) + theme(
        legend.position='bottom'
    )
    cur_plot_list[[cur_group]] <- p

  
    for(cur_gene in gene_list){

        # plot the clusters in space
        p <- plotSpatialFeature(
                plot_sfe, cur_gene, 
                colGeometryName = "cellSeg",
                bbox = bbox
            ) + 
            ggtitle(cur_gene) + 
            labs(fill="") + theme(
                panel.border = element_rect(linewidth=0.5, color='black', fill=NA),
                plot.title = element_text(hjust=0.5, face='bold'),
                legend.position='bottom'
            )
            
        cur_plot_list[[cur_gene]] <- p
    }
    plot_list[[cur_group]] <- cur_plot_list

}
names(bbox_df) <- c('xmin', 'ymin', 'xmax', 'ymax')

pdf(paste0(cur_sample, '/figures/', 'degs_top5_micromets_zoom.pdf'), width=15, height=5)

    for(cur_plots in plot_list){
        patch <- wrap_plots(cur_plots, ncol=6)
        print(patch)
    }

dev.off()



```


Pie charts showing % of cells expressing markers 

```{r eval=FALSE}

# gene list for micromet genes:
gene_list <- c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A')

# expresion matrix:
X <- counts(sfe_tumor)


groups <- levels(sfe_tumor$spatial_community)
plot_df <- data.frame()
mxd4_df <- data.frame() 
for(cur_group in c(groups, 'all')){
    print(cur_group)

    # get current exp
    if(cur_group == 'all'){
        cur_X <- X
    } else{
        cur_X <- X[,sfe_tumor$spatial_community %in% cur_group]
    }
    cur_binary <- cur_X; cur_binary[cur_binary > 1 ] <- 1

    # what % express any of these markers?
    sum(colSums(cur_binary[gene_list,]) > 0) / ncol(cur_X) * 100 
    n_list <- c()
    for(i in 0:length(gene_list)){
        n_list <- c(n_list,
            sum(colSums(cur_binary[gene_list,]) == i)
        )
    }
    
    # what % express MXD4?
    cur_mxd4 <- sum(cur_binary["MXD4", ]) / ncol(cur_X)
    mxd4_df <- rbind(
        mxd4_df, 
        data.frame(
            proportion = c(cur_mxd4, 1-cur_mxd4),
            group = c(cur_group, cur_group),
            pos = c('MXD4+', 'MXD4-')
    ))


    cur_df <- data.frame(
        'n_cells' = n_list,
        'proportion' = n_list / ncol(cur_X),
        'n_genes' = 0:length(gene_list),
        'group' = cur_group
    )

    plot_df <- rbind(plot_df, cur_df)

}

sum(is.na(plot_df))
sum(is.na(plot_df$n_genes))

unique(plot_df$n_genes)

# set factor levels
plot_df$n_genes <- factor(
    as.numeric(plot_df$n_genes),
    levels = 0:length(gene_list)
)
sum(is.na(plot_df$n_genes))


#plot_df$n_genes <- as.numeric(plot_df$n_genes)

plot_df$group <- factor(
    as.character(plot_df$group),
    levels = c(levels(sfe_tumor$spatial_community), 'all')
)

# Get the positions
df2 <- plot_df %>% group_by(group) %>%
  mutate(csum = rev(cumsum(rev(proportion))), 
         pos = proportion/2 + lead(csum, 1),
         pos = if_else(is.na(pos), proportion/2, pos))



# make pie charts
p <- plot_df %>% subset(proportion > 0) %>%
    ggplot(aes(x = "", y = proportion, fill=n_genes)) + 
    geom_col(color='black') + 
    ggrepel::geom_text_repel(
        data = subset(df2, proportion > 0.05),
        aes( y = pos,
            label = paste0(round(proportion, digits=2)*100, '%')), 
            size=3, nudge_x = 1, show.legend=FALSE) + 
    coord_polar(theta="y") + 
    #viridis::scale_fill_viridis() + 
    scale_fill_brewer() +
    theme_void()


patch <- p + facet_wrap(~group, ncol=10)

pdf(paste0(cur_sample, '/figures/', 'micromet_pie.pdf'), width=18, height=6)
patch 
dev.off()


# plot the mxd4% 
group_order <- mxd4_df %>% subset(pos == 'MXD4+') %>% arrange(desc(proportion)) %>% .$group


mxd4_df$group <- factor(as.character(mxd4_df$group), levels=group_order)

p <- mxd4_df %>% 
    ggplot(aes(x=group, y=proportion, fill=pos)) +
    geom_bar(position='fill', stat='identity') +
    ylab('Proportion') + 
    xlab('Met group') + scale_fill_manual(values=c('grey90', 'grey70')) + 
    labs(fill='') + RotatedAxis()

pdf(paste0(cur_sample, '/figures/', 'mxd4_proportions.pdf'), width=8, height=2)
p
dev.off()


```

How many tumor cells are MXD4+ ?

```{r eval=FALSE}

"MXD4" %in% rownames(sfe_tumor)

# expresion matrix:
X <- counts(sfe_tumor)
binary <- X; binary[binary > 1 ] <- 1



```

MicroMet FindMarkers (?)

```{r eval=FALSE}

# plot vln plots with seurat
X <- logcounts(sfe_tumor)
seurat_obj <- CreateSeuratObject(
    counts=X, meta.data=as.data.frame(colData(sfe_tumor))
)
Idents(seurat_obj) <- seurat_obj$spatial_community

# loop through each signature:
markers <- FindAllMarkers(
    seurat_obj,
    slot = 'counts',
    test.use = 'wilcox',
    only.pos=TRUE
) %>% subset(p_val_adj <= 0.05 & avg_log2FC > 1)

head(markers)
table(markers$cluster)
length(unique(markers$gene))


```

ComplexHeatmap of micromets (?)

```{r eval=FALSE}

library(ComplexHeatmap)

genes_plot <- markers %>% group_by(cluster) %>% slice_max(n=5, order_by=avg_log2FC) %>%  .$gene %>% unique()

# calculate pseudo-bulk expression of tumor cells per micromet
X <- counts(sfe_tumor) 
groups <- levels(sfe_tumor$spatial_community)

pb_df <- data.frame()
for(cur_group in groups){
    cur_pb <- rowSums(X[,sfe_tumor$spatial_community == cur_group])
    cur_df <- data.frame(gene = names(cur_pb), count = as.numeric(cur_pb), group = cur_group)
    pb_df <- rbind(pb_df, cur_df)
}
X_pb <- pb_df %>% tidyr::pivot_wider(names_from=gene, values_from=count) %>% as.data.frame()
rownames(X_pb) <- X_pb[,1]
X_pb <- X_pb[,-1]
X_pb <- as.matrix(X_pb)

# normalize
X_norm <- t(apply(X_pb, 1, function(x){
     y <- (x) / sum(x) * 10000
    log1p(y + 1)
}))

plot_mat <- base::scale(X_norm[,genes_plot])

pdf(paste0(cur_sample, '/figures/', 'micromet_heatmap.pdf'), width=6, height=15)

Heatmap(t(plot_mat), cluster_rows=FALSE, cluster_columns=FALSE)

dev.off()

```  