
Load the 5 datasets together


```{r eval=FALSE}


library(Voyager)
library(SpatialFeatureExperiment)
library(rjson)
library(Matrix)
library(SFEData)
library(RBioFormats)


library(SingleCellExperiment)
library(SpatialExperiment)
library(ggplot2)
library(stringr)
library(scater) 
library(scuttle)
library(BiocParallel)
library(BiocSingular)
library(bluster)
library(scran)
library(patchwork)
library(cowplot)
theme_set(theme_cowplot())
library(Seurat)
library(tidyverse)

set.seed(22)

setwd('~/collab/Valiente_2024/individual/')
data_dir <- 'data/'
fig_dir <- 'figures/'


# load the 5 samples:

sample_list <- c('134990_left', '134990_right', '10010900', '10032256', '10058583')
sfe_list <- c()
X_list <- list() 
colData_list <- list()
rowData_list <- list() 
for(cur_sample in sample_list){
    print(cur_sample)
    x <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))
    sfe_list[[cur_sample]] <-x
    
    cur_X <- counts(x)
    colnames(cur_X) <- paste0(cur_sample, '-', colnames(cur_X))
    X_list[[cur_sample]] <- cur_X

    cur_colData <- colData(x)
    rownames(cur_colData) <- paste0(cur_sample, '-', rownames(cur_colData))
    colData_list[[cur_sample]] <- cur_colData

    rowData_list[[cur_sample]] <- rowData(x)
}

columns_keep <- Reduce(intersect, lapply(sfe_list, function(x){colnames(colData(x))}))


new_colData <- Reduce(rbind, lapply(sample_list, function(cur_sample){

    cur_colData <- colData_list[[cur_sample]]

    cur_colData[,columns_keep]

}))
new_X <- Reduce(cbind, X_list)

# create new sce object
sce <- SingleCellExperiment(
    new_X
)
names(assays(sce)) <- 'counts'

colData(sce) <- new_colData
rowData(sce) <- rowData(sfe_list[[1]])

# fix the micromet cluster:
colData(sce)$spatial_community <- paste0(as.character(colData(sce)$SampleID), '-', as.character(colData(sce)$spatial_community))
table(sce$spatial_community)



```


Pie charts showing % of cells expressing markers 

```{r eval=FALSE}

# gene list for micromet genes:
gene_list <- c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A')

# expresion matrix:
sfe_tumor <- sce[,sce$cell_group == 'Tumor']
mm_exclude <- names(which(table(sfe_tumor$spatial_community) == 1))
sfe_tumor <- sfe_tumor[,!sfe_tumor$spatial_community %in% mm_exclude]


X <- counts(sfe_tumor)


groups <- unique(sce$spatial_community)
plot_df <- data.frame()
mxd4_df <- data.frame() 
for(cur_group in c('all')){
#for(cur_group in c(groups, 'all')){

    print(cur_group)

    # get current exp
    if(cur_group == 'all'){
        cur_X <- X
    } else{
        cur_X <- X[,sfe_tumor$spatial_community %in% cur_group]
    }
    cur_binary <- as(cur_X, 'dgCMatrix'); cur_binary[cur_binary > 1 ] <- 1
    ncells <- ncol(cur_X)

    # what % express any of these markers?
    #sum(colSums(cur_binary[gene_list,]) > 0) / ncol(cur_X) * 100 
    n_list <- c()
    any_list <- c()
    for(i in 1:length(gene_list)){
        n_list <- c(n_list,
            sum(colSums(cur_binary[gene_list,]) == i)
        )
    }

    # what % express any marker `
    any_list <- c(any_list, sum(colSums(cur_binary[gene_list,]) > 0))

    
    # what % express MXD4?
    cur_mxd4 <- sum(cur_binary["MXD4", ]) / ncells
    mxd4_df <- rbind(
        mxd4_df, 
        data.frame(
            proportion = c(cur_mxd4, 1-cur_mxd4),
            group = c(cur_group, cur_group),
            pos = c('MXD4+', 'MXD4-'),
            n_cells = c(ncells, ncells)
    ))

    



    cur_df <- data.frame(
        'n_cells' = n_list,
        'any_cells' = any_list,
        'proportion' = n_list / ncol(cur_X),
        'n_genes' = 1:length(gene_list),
        'group' = cur_group,
        'n_total' = ncells
    )

    plot_df <- rbind(plot_df, cur_df)

}

sum(is.na(plot_df))
sum(is.na(plot_df$n_genes))

unique(plot_df$n_genes)


# set factor levels
plot_df$n_genes <- factor(
    as.numeric(plot_df$n_genes),
    levels = 0:length(gene_list)
)
sum(is.na(plot_df$n_genes))


#plot_df$n_genes <- as.numeric(plot_df$n_genes)


# Get the positions
df2 <- plot_df %>% subset(group == 'all') %>%
  mutate(csum = rev(cumsum(rev(proportion))), 
         pos = proportion/2 + lead(csum, 1),
         pos = if_else(is.na(pos), proportion/2, pos))



# make pie charts
p <- plot_df %>% subset(group == 'all') %>%
    ggplot(aes(x = "", y = proportion, fill=n_genes)) + 
    geom_col(color='black') + 
    # ggrepel::geom_text_repel(
    #     data = df2,
    #     aes( y = pos,
    #         label = paste0(round(proportion, digits=2)*100, '%')), 
    #         size=3, nudge_x = 1, show.legend=FALSE) + 
    coord_polar(theta="y") + 
    #viridis::scale_fill_viridis() + 
    scale_fill_brewer() +
    theme_void()

pdf(paste0('micromet_pie_all.pdf'), width=5, height=5)
p
dev.off()


any_df <- plot_df[1,c('any_cells', 'n_total')]
any_df <- rbind(
    any_df, data.frame(
        'any_cells' = any_df$n_total - any_df$any_cells,
        'n_total' = any_df$n_total
    )
)
any_df$group <- c('expressed', 'not expressed')
any_df$proportion <- any_df$any_cells / any_df$n_total

# Get the positions
df2 <- any_df %>%
  mutate(csum = rev(cumsum(rev(proportion))), 
         pos = proportion/2 + lead(csum, 1),
         pos = if_else(is.na(pos), proportion/2, pos))



# make pie charts
p <- any_df %>%
    ggplot(aes(x = "", y = proportion, fill=group)) + 
    geom_col(color='black') + 
    # ggrepel::geom_text_repel(
    #     data = df2,
    #     aes( y = pos,
    #         label = paste0(round(proportion, digits=2)*100, '%')), 
    #         size=3, nudge_x = 1, show.legend=FALSE) + 
    coord_polar(theta="y") + 
    #viridis::scale_fill_viridis() + 
    scale_fill_manual(values=c('grey70', 'grey95')) +
    theme_void()

pdf(paste0('micromet_pie_any.pdf'), width=5, height=5)
p
dev.off()


tmp <- na.omit(plot_df)

cor(tmp$proportion, tmp$n_total)


```


# Analyze signatures

```{r eval=FALSE}

#-----------------------------------------------------------------
# Calculate Signatures (only in Tumor cells!!!)
#-----------------------------------------------------------------

genes_signature = list(
    'micromets'= c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A'),
    'dormancy'= c('NDRG1', 'DYRK1A', 'ZEB2', 'THBS1', 'BMPR2', 'TGFBR3', 'AXL', 'BHLHE41', 'RHEB', 'NR2F1'),
    'stress_general'= c('HIF1A', 'H2AZ1', 'NFE2L2', 'HSP90AA1', 'ATF6', 'ATM', 'ATR', 'NRF2', 'ATF3'),
    'proliferation'= c('CDK4', 'PCNA', 'MKI67', 'TOP2A', 'CCNB2'),
    'stress_ATF6' = c('ATF6',	'HSP90AA1',	'XBP1',	'CIP2A'),
    'stress_DNA_damage' = c('H2AZ1',	'ATM',	'ATR'),
    'dormancy_expanded' = c('RHEB',	'NR2F1',	'SOX9',	'AXL',	'ATF6',	'BMPR2',	'TGFBR3',	'RPS6KA5',	'NDRG1',	'BHLHE41',	'MTOR',	'DYRK1A',	'ZEB2',	'MAPK14',	'THBS1'),
    'autophagy' = c('SQSTM1',	'BECN1',	'MAP1LC3B',	'CAV1',	'RPTOR',	'LAMP3'),
    'chromosomal_instability' = c('PCNA',	'UBE2C',	'CCNB2',	'TOP2A')
)

# adjust the signatures to only include genes in the dataset 
for(sig in names(genes_signature)){
    cur_genes <- genes_signature[[sig]]
    genes_signature[[sig]] <- cur_genes[cur_genes %in% rowData(sfe_list[[1]])$Symbol]
}

# set up tumor dataset
sfe_tumor <- sce[,sce$cell_group == 'Tumor']
mm_exclude <- names(which(table(sfe_tumor$spatial_community) == 1))
sfe_tumor <- sfe_tumor[,!sfe_tumor$spatial_community %in% mm_exclude]
sfe_tumor <- logNormCounts(sfe_tumor)

meta <- colData(sfe_tumor)

test <- UCell::ScoreSignatures_UCell(
    as(counts(sfe_tumor), 'dgCMatrix'), 
    features=genes_signature
)
colnames(test) <- names(genes_signature)
colData(sfe_tumor) <- cbind(colData(sfe_tumor), test)

#--------------------------------------------------------------------#
# micromet-level signature plots (raincloud plot)
#--------------------------------------------------------------------#

plot_df <- data.frame()
for(cur_sig in names(genes_signature)){

    cur_df <- as.data.frame(colData(sfe_tumor)[,c('spatial_community', cur_sig, 'SampleID')])
    colnames(cur_df)[2] <- 'score'
    cur_df$signature <- cur_sig 
    plot_df <- rbind(plot_df, cur_df)
}

# compute averages
plot_df <- plot_df %>% 
    group_by(spatial_community, SampleID, signature) %>%
    summarise(value = mean(score), ncells = n())

signatures <- names(genes_signature)
sig1 <- 'micromets'

plot_list <- list()
cor_df <- data.frame()
for(sig1 in signatures){

    for(sig2 in signatures){

        print(paste0(sig1, sig2))
        cur_df <- subset(plot_df, signature %in% c(sig1, sig2))
        cur_df <- cur_df %>% 
                    tidyr::pivot_wider(names_from=signature, values_from=value)
                cur_df <- cur_df[,c('spatial_community', sig1, sig2)]
                colnames(cur_df) <- c('spatial_community', 'sig1', 'sig2')

        if(sig1 == sig2){next}
        if(paste0(sig2, '_', sig1) %in% names(plot_list)){
            plot_list[[paste0(sig1, '_', sig2)]] <- ggplot() + theme_void()
            next
        }

        cur_cor <- cor(cur_df$sig1, cur_df$sig2)
        cor_df <- rbind(cor_df, data.frame(cor=cur_cor, sig1=sig1, sig2=sig2))

        p <- cur_df %>% 
            ggplot(aes(x=sig1, y=sig2, color=spatial_community)) + 
            geom_point() + 
            geom_smooth(
                inherit.aes=FALSE,
                aes(x=sig1, y=sig2),
                method='lm'
            ) + 
            ggpubr::stat_cor() + 
            xlab(sig1) + ylab(sig2) + theme(
                axis.line.y = element_blank(),
                axis.line.x = element_blank(),
                panel.border = element_rect(linewidth=0.5, fill=NA, color='black')
            )

        plot_list[[paste0(sig1, '_', sig2)]] <- p
    }
}


cor_df$sig1 <- factor(as.character(cor_df$sig1), levels=signatures)
cor_df$sig2 <- factor(as.character(cor_df$sig2), levels=signatures)

p <- cor_df %>% 
    ggplot(aes(x=sig1, y=sig2, fill=cor)) + 
    geom_tile() + 
    coord_fixed() + RotatedAxis() + 
    scale_fill_gradient2(high='red', mid='white', low='blue', limits=c(-1,1)) + 
    geom_text(
        data = subset(cor_df, abs(cor) > 0.5), 
        aes(label=round(cor, 2))
    ) + xlab('') + ylab('')


pdf(paste0('signatures_compare_heatmap_all.pdf'), width=6, height=6)
p
dev.off()



#-----------------------------------------------------------------
# Correlation between the micromet signatures vs size
#-----------------------------------------------------------------

p <- plot_df %>%
    ggplot(aes(x=ncells, y=value)) + 
    geom_point() + 
    geom_smooth(method='lm') + 
    ggpubr::stat_cor()



patch <- p + facet_grid(SampleID~signature, scales='free')

pdf(paste0('signatures_vs_size.pdf'), width=15, height=5)
patch 
dev.off()


#-----------------------------------------------------------------
# Pairwise correlation of micromets based on their signatures.
#
# Also account for the distance between MMs here?
#-----------------------------------------------------------------


# # compute the cluster centroids (for plotting labels)
centroid_df <- colData(sfe_tumor) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))


groups <- unique(plot_df$spatial_community)
cor_df <- data.frame()
for(group1 in groups){
    for(group2 in groups){
        if(group1 == group2){next}

        compare <- paste0(group1, '_', group2)
        compare_rev <- paste0(group2, '_', group1)
        if(compare_rev %in% cor_df$compare){next}

        print(paste0(group1, ' vs. ', group2))
        # correlation
        cur_df1 <- plot_df %>% subset(spatial_community == group1)
        cur_df2 <- plot_df %>% subset(spatial_community == group2)
        result <- cor.test(cur_df1$value, cur_df2$value)

        # distance? 
        coords <- subset(centroid_df, spatial_community %in% c(group1, group2)) %>% 
            select(-spatial_community)
        cur_dist <- as.numeric(dist(as.matrix(coords), method='euclidean'))

        cur_result <- data.frame(
            group1 = group1, 
            group2 = group2,
            cor = as.numeric(result$estimate),
            pval = as.numeric(result$p.value),
            dist = cur_dist, 
            compare = compare
        )

        cor_df <- rbind(cor_df, cur_result)

    }
}

cor_df$sample1 <- do.call(rbind, str_split(cor_df$group1, '-'))[,1]
cor_df$sample2 <- do.call(rbind, str_split(cor_df$group2, '-'))[,1]

cor_df$fdr <- p.adjust(cor_df$pval, 'fdr')
table(cor_df$fdr <= 0.05)

p <- cor_df %>% subset(fdr <= 0.01) %>%
    ggplot(aes(x=group1, y=group2, fill=cor)) +
    geom_tile() + 
    coord_fixed() + RotatedAxis() + 
    scale_fill_gradient2(high='red', mid='white', low='blue', limits=c(-1,1)) + theme_void()


tmp <- cor_df %>% select(c(group1, sample1)) %>% distinct()
cb_df <- data.frame(
    v1 = 1,
    v2 = tmp$group1,
    sample = tmp$sample1
)

cb1 <- cb_df %>% 
    ggplot(aes(y=v1,x=v2, fill=sample)) + 
    geom_bar(position='stack', stat='identity') +
    theme_void()



tmp <- cor_df %>% select(c(group2, sample2)) %>% distinct()
cb_df <- data.frame(
    v1 = 1,
    v2 = tmp$group2,
    sample = tmp$sample2
)

# cb_df <- data.frame(
#     v1 = 1,
#     v2 = c('a', 'b', 'c', 'd'),
#     sample = c('aa', 'aa', 'bb', 'bb')
# )

cb2 <- cb_df %>% 
    ggplot(aes(x=v1, y=v2, fill=sample)) + 
    geom_bar(stat='identity', position='stack') + theme_void()

layout <- "
#A
BC
"

patch <- cb1 + cb2 + p + plot_layout(design = layout, widths=c(1,50), heights=c(1,50), guides='collect')


pdf(paste0('micromet_correlations2.pdf'), width=11, height=10)
patch
dev.off()



cor_df$Sample1 <- do.call(rbind, str_split(cor_df$group1, '-'))[,1]
cor_df$Sample2 <- do.call(rbind, str_split(cor_df$group2, '-'))[,1]

# save the table:
cor_df %>% 
    select(-dist) %>% 
    write.csv('micromet_correlations.csv', quote=F, row.names=F)



#-----------------------------------------------------------------
# Pairwise correlation of micromets based on expression of all genes
#
# Also account for the distance between MMs here?
#-----------------------------------------------------------------


# # compute the cluster centroids
centroid_df <- colData(sfe_tumor) %>%
    as.data.frame() %>%
    group_by(spatial_community) %>% 
    summarise(x = mean(x_centroid), y = mean(y_centroid))

# calculate pseudo-bulk expression of tumor cells per micromet
X <- counts(sfe_tumor) 
groups <- unique(sfe_tumor$spatial_community)

pb_df <- data.frame()
for(cur_group in groups){
    print(cur_group)
    cur_pb <- rowSums(X[,sfe_tumor$spatial_community == cur_group])
    cur_df <- data.frame(gene = names(cur_pb), count = as.numeric(cur_pb), group = cur_group)
    pb_df <- rbind(pb_df, cur_df)
}

X_pb <- pb_df %>% tidyr::pivot_wider(names_from=gene, values_from=count) %>% as.data.frame()
rownames(X_pb) <- X_pb[,1]
X_pb <- X_pb[,-1]
X_pb <- as.matrix(X_pb)

# normalize
X_norm <- t(apply(X_pb, 1, function(x){
     y <- (x) / sum(x) * 10000
    log1p(y + 1)
}))


groups <- unique(sfe_tumor$spatial_community)
cor_df <- data.frame()
for(group1 in groups){
    for(group2 in groups){
        if(group1 == group2){next}

        compare <- paste0(group1, '_', group2)
        compare_rev <- paste0(group2, '_', group1)
       # if(compare_rev %in% cor_df$compare){next}

        print(paste0(group1, ' vs. ', group2))
        # correlation
        x <- X_norm[group1,]
        y <- X_norm[group2,]
        result <- cor.test(x, y)

        # distance? 
        coords <- subset(centroid_df, spatial_community %in% c(group1, group2)) %>% 
            select(-spatial_community)
        cur_dist <- as.numeric(dist(as.matrix(coords), method='euclidean'))

        cur_result <- data.frame(
            group1 = group1, 
            group2 = group2,
            cor = as.numeric(result$estimate),
            pval = as.numeric(result$p.value),
            dist = cur_dist, 
            compare = compare
        )

        cor_df <- rbind(cor_df, cur_result)

    }
}



cor_df$sample1 <- do.call(rbind, str_split(cor_df$group1, '-'))[,1]
cor_df$sample2 <- do.call(rbind, str_split(cor_df$group2, '-'))[,1]

cor_df$fdr <- p.adjust(cor_df$pval, 'fdr')
table(cor_df$fdr <= 0.001)

p <- cor_df %>% subset(fdr <= 0.01) %>%
    ggplot(aes(x=group1, y=group2, fill=cor)) +
    geom_tile() + 
    coord_fixed() + RotatedAxis() + 
    scale_fill_gradient2(high='red', mid='white', low='blue', limits=c(-1,1)) + theme_void()


tmp <- cor_df %>% select(c(group1, sample1)) %>% distinct()
cb_df <- data.frame(
    v1 = 1,
    v2 = tmp$group1,
    sample = tmp$sample1
)

cb1 <- cb_df %>% 
    ggplot(aes(y=v1,x=v2, fill=sample)) + 
    geom_bar(position='stack', stat='identity') +
    theme_void()



tmp <- cor_df %>% select(c(group2, sample2)) %>% distinct()
cb_df <- data.frame(
    v1 = 1,
    v2 = tmp$group2,
    sample = tmp$sample2
)

cb2 <- cb_df %>% 
    ggplot(aes(x=v1, y=v2, fill=sample)) + 
    geom_bar(stat='identity', position='stack') + theme_void()

layout <- "
#A
BC
"

patch <- cb1 + cb2 + p + plot_layout(design = layout, widths=c(1,50), heights=c(1,50), guides='collect')


pdf(paste0('micromet_correlations_expression.pdf'), width=11, height=10)
patch
dev.off()



cor_df$Sample1 <- do.call(rbind, str_split(cor_df$group1, '-'))[,1]
cor_df$Sample2 <- do.call(rbind, str_split(cor_df$group2, '-'))[,1]

# save the table:
cor_df %>% 
    select(-dist) %>% 
    write.csv('micromet_correlations_expression.csv', quote=F, row.names=F)





#-----------------------------------------------------------------
# Correlation between mxd4 expression and micromet size
#-----------------------------------------------------------------

# add mxd4 as it's own "signature"
cur_exp <- logcounts(sfe_tumor)["MXD4",]
plot_df <- data.frame(
    spatial_community = sfe_tumor$spatial_community,
    score = cur_exp,
    SampleID = sfe_tumor$SampleID,
    signature = 'MXD4'
)

# compute averages
plot_df <- plot_df %>% 
    group_by(spatial_community, SampleID, signature) %>%
    summarise(value = mean(score), ncells = n())

cor.test(plot_df$ncells, plot_df$value)


p <- plot_df %>% 
    ggplot(aes(x=value, y=ncells, color=SampleID)) + 
    geom_point() + 
    geom_smooth(
        data = plot_df, inherit.aes=FALSE,
        mapping=aes(x=value, y=ncells),
        method='lm'
    )+ 
    ggpubr::stat_cor(
         data = plot_df, inherit.aes=FALSE,
        mapping=aes(x=value, y=ncells),
    ) + xlab('MXD4 Expression') + ylab('N cells')

patch <- p + facet_wrap(~SampleID, scales='free', ncol=5)

pdf(paste0('mxd4_vs_size.pdf'), width=5, height=4)
p
dev.off()

pdf(paste0('mxd4_vs_size_split.pdf'), width=15, height=3)
p + facet_wrap(~SampleID, scales='free', ncol=5)
dev.off()




p <- plot_df %>% 
    ggplot(aes(x=value, y=ncells, color=SampleID)) + 
    geom_point() + 
    geom_smooth(
        data = plot_df, inherit.aes=FALSE,
        mapping=aes(x=value, y=ncells),
        method='lm'
    ) #+
    #scale_y_continuous(breaks = seq(0,1500,300), limits = c(0,1500)) +
   # ggbreak::scale_y_break(c(1000,1400), space = 0.1, ticklabels = c(seq(0,1500,300), 1400)) + 
    # theme(
    #     axis.title.y.right = element_blank(),
    #     axis.ticks.y.right = element_blank(),
    #     axis.line.y.right = element_blank(),
    #     axis.text.y.right = element_blank()
    # )


#+ xlab('MXD4 Expression') + ylab('N cells') + 

#patch <- p + facet_wrap(~SampleID, scales='free', ncol=5)



pdf(paste0('mxd4_vs_size_nolabel.pdf'), width=5, height=4)
p
dev.off()

pdf(paste0('mxd4_vs_size_split_nolabel.pdf'), width=15, height=3)
p + facet_wrap(~SampleID, scales='free', ncol=5)
dev.off()


```  

Proportion plots but fix the x axis labels:

```{r eval=FALSE}


#-----------------------------------------------------------------
# Plot cell-type compositions of the micromets:
#-----------------------------------------------------------------

# define color palettes
cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
names(cp) <- c("Immune", 'Tumor', "Vasculature",  'other' )


for(cur_sample in sample_list){

    print(cur_sample)
    sfe_sub <- sfe_list[[cur_sample]]

    plot_df <- table(sfe_sub$spatial_community, sfe_sub$cell_group) %>% 
        as.data.frame() %>%
        reshape2::melt()
    colnames(plot_df)[1:2] <- c('spatial_community', 'cell_group')



    p <- plot_df %>% 
        ggplot(aes(x=spatial_community, y=value, fill=cell_group)) +
        geom_bar(position='fill', stat='identity') +
        ylab('Proportion') + 
        xlab('Met group') +
        scale_fill_manual(values=cp) + Seurat::RotatedAxis()

    w <-  round((length(levels(sfe_sub$spatial_community)) / 5) + 1)

    pdf(paste0(cur_sample, '/figures/', 'spatial_community_proportions.pdf'), width=w, height=4)
    print(p)
    dev.off()

    # make another plot but order by tumor fraction
    tmp <- plot_df %>% 
        group_by(spatial_community) %>% 
        summarise(total = sum(value))

    tmp2 <- plot_df %>% subset(cell_group == 'Tumor')
    frac <- tmp2$value / tmp$total
    names(frac) <- as.character(tmp$spatial_community)

    plot_df$spatial_community <- factor(
        as.character(plot_df$spatial_community),
        levels = as.character(names(frac)[order(-frac)])
    )

    p <- plot_df %>% 
        ggplot(aes(x=spatial_community, y=value, fill=cell_group)) +
        geom_bar(position='fill', stat='identity') +
        ylab('Proportion') + 
        xlab('Met group') +
        scale_fill_manual(values=cp) + Seurat::RotatedAxis()


    pdf(paste0(cur_sample, '/figures/', 'spatial_community_proportions_ordered.pdf'), width=w, height=4)
    print(p)
    dev.off()

}

```



Micromet Zoom Plots 

```{r eval=FALSE}


genes_signature = list(
    'micromets'= c('SOX2', 'BCL2', 'MXD4', 'CDKN1B', 'CDKN1A'),
    'dormancy'= c('NDRG1', 'DYRK1A', 'ZEB2', 'THBS1', 'BMPR2', 'TGFBR3', 'AXL', 'BHLHE41', 'RHEB', 'NR2F1'),
    'stress_general'= c('HIF1A', 'H2AZ1', 'NFE2L2', 'HSP90AA1', 'ATF6', 'ATM', 'ATR', 'NRF2', 'ATF3'),
    'proliferation'= c('CDK4', 'PCNA', 'MKI67', 'TOP2A', 'CCNB2'),
    'stress_ATF6' = c('ATF6',	'HSP90AA1',	'XBP1',	'CIP2A'),
    'stress_DNA_damage' = c('H2AZ1',	'ATM',	'ATR'),
    'dormancy_expanded' = c('RHEB',	'NR2F1',	'SOX9',	'AXL',	'ATF6',	'BMPR2',	'TGFBR3',	'RPS6KA5',	'NDRG1',	'BHLHE41',	'MTOR',	'DYRK1A',	'ZEB2',	'MAPK14',	'THBS1'),
    'autophagy' = c('SQSTM1',	'BECN1',	'MAP1LC3B',	'CAV1',	'RPTOR',	'LAMP3'),
    'chromosomal_instability' = c('PCNA',	'UBE2C',	'CCNB2',	'TOP2A')
)

#-------------------------------------------------------#
# Zoom plots of the individual micromet genes
#-------------------------------------------------------#

for(cur_sample in sample_list){
    
    sfe <- sfe_list[[cur_sample]]
    plot_sfe <- sfe[,sfe$cell_group == 'Tumor']
    gene_list <- genes_signature$micromets
    plot_groups <- levels(sfe$spatial_community)

    # define color palettes
    cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
    names(cp) <- c("Immune", 'Tumor', "Vasculature",  'other' )

    # bounding box offset
    offset <- 25

    bbox_df <- data.frame()
    plot_list <- list()
    cluster_plot_list <- list()
    for(cur_group in plot_groups){

        print(cur_group)
        cur_plot_list <- list()

        # define the bbox for this group
        cur_df <- colData(sfe) %>% as.data.frame() %>%
            subset(spatial_community == cur_group)

        # current plotting group
        colData(sfe)$plot_group <- ifelse(
            colData(sfe)$spatial_community == cur_group & colData(sfe)$spatial_community != 'other',
            colData(sfe)$cell_group, 'other'
        )
        colData(sfe)$plot_group <- factor(
            as.character(colData(sfe)$plot_group),
            levels = c("Immune", 'Tumor', "Vasculature",  'other' )
        )
 
        # define bounding box
        bbox <- c(
            xmin = min(cur_df$x_centroid) - offset,
            ymin = min(cur_df$y_centroid) - offset,
            xmax = max(cur_df$x_centroid) + offset,
            ymax = max(cur_df$y_centroid) + offset
        )

        bbox_df <- rbind(bbox_df, bbox)

        p <- plotSpatialFeature(
            sfe, 'plot_group', 
            colGeometryName = "cellSeg",
            bbox = bbox
        ) +
        ggtitle(paste0("Micromet ", cur_group)) + 
        labs(fill="") +
        scale_fill_manual(values=cp) + theme(
            plot.title = element_text(face='bold', hjust=0.5),
            panel.border=element_rect(fill=NA, linewidth=0.5, color='black')
        ) + theme(
            legend.position='bottom'
        )
        cur_plot_list[[cur_group]] <- p

    
        for(cur_gene in gene_list){

            # plot the clusters in space
            p <- plotSpatialFeature(
                    plot_sfe, cur_gene, 
                    colGeometryName = "cellSeg",
                    bbox = bbox
                ) + 
                ggtitle(cur_gene) + 
                labs(fill="") + theme(
                    panel.border = element_rect(linewidth=0.5, color='black', fill=NA),
                    plot.title = element_text(hjust=0.5, face='bold'),
                    legend.position='bottom'
                )
                
            cur_plot_list[[cur_gene]] <- p
        }
        plot_list[[cur_group]] <- cur_plot_list

    }
    names(bbox_df) <- c('xmin', 'ymin', 'xmax', 'ymax')

    pdf(paste0(cur_sample, '/figures/', 'signature_micromets_zoom.pdf'), width=15, height=5)

        for(cur_plots in plot_list){
            patch <- wrap_plots(cur_plots, ncol=6)
            print(patch)
        }

    dev.off()
}


```

UMAP plots colored by cell group 

```{r eval=FALSE}

# define color palettes
    cp <- c(dittoSeq::dittoColors()[1:3], 'lightgrey')
    names(cp) <- c("Immune", 'Tumor', "Vasculature",  'other' )


for(cur_sample in sample_list){
    
    sfe <- sfe_list[[cur_sample]]
    
    # plot the UMAP colored by cluster
    p <- plotUMAP(sfe, colour_by = 'cell_group') + scale_color_manual(values=cp)
    pdf(paste0(cur_sample, '/figures/', 'umap_celltypes.pdf'), width=7, height=7)
    print(p)
    dev.off()

}

```


Size of each micromet - make a table and make a box plot 

```{r eval=FALSE}

plot_df <- data.frame()
for(cur_sample in sample_list){
    
    sfe <- sfe_list[[cur_sample]]
    sfe <- sfe[,sfe$cell_group == 'Tumor']
    
    # micromet sizes 
    x <- table(sfe$spatial_community)

    plot_df <- rbind(plot_df, data.frame(
        micromet = names(x),
        ncells = as.numeric(x),
        sample = cur_sample
    ))

}

p <- plot_df %>% 
    ggplot(aes(y=ncells, x=sample, fill=sample)) + 
    geom_boxplot(outlier.shape=NA, width=0.6) + 
    geom_jitter(width=0.2, size=1) + 
    ylab('N cells') +
    RotatedAxis() + NoLegend()


pdf(paste0('micromet_sizes.pdf'), width=4, height=7)
print(p)
dev.off()

p <- plot_df %>% 
    ggplot(aes(y=ncells, x="all")) + 
    geom_boxplot(outlier.shape=NA, width=0.6, fill='grey80') + 
    geom_jitter(width=0.2, size=1) + 
    ylab('N cells') + xlab('') +
    RotatedAxis() + NoLegend()

pdf(paste0('micromet_sizes_all.pdf'), width=2, height=3)
print(p)
dev.off()

# number of mircomets per sample:
tmp <- table(plot_df$sample)
df <- data.frame(
    value = as.numeric(tmp),
    sample = names(tmp)
) %>% arrange(desc(value)) 
df$sample <- factor(as.character(df$sample), levels =as.character(df$sample))

p <- df %>% 
    ggplot(aes(x=sample, y=value)) + 
    geom_bar(stat='identity', fill='grey65') + RotatedAxis() + 
    xlab('') + ylab('N micromets') 

pdf(paste0('micromet_nums.pdf'), width=4, height=7)
print(p)
dev.off()


p <- p + geom_text(aes(label=value), vjust=-0.2)

pdf(paste0('micromet_nums_labeled.pdf'), width=4, height=7)
print(p)
dev.off()



```

Calculations 

# make an excel sheet with this info 

```{r eval=FALSE}

# number of cancer cells in total 
print(dim(sfe_tumor)) # 46,584 tumor cells total 

# number of tumor cells total per patient
table(sfe_tumor$SampleID)
10010900     10032256     10058583  134990_left 134990_right 
    5955         7610        28389         2169         2461 

# mean number of cancer cells p
mean_list <- c()
for(cur_sample in sample_list){
    cur_df <- colData(sfe_tumor) %>% subset(SampleID == cur_sample)
    mm_sizes <- table(cur_df$spatial_community)
    mean_list <- c(mean_list, mean(mm_sizes))
}
names(mean_list) <- sample_list 
mean_list
#  134990_left 134990_right     10010900     10032256     10058583 
#     74.79310     41.71186     78.35526     69.18182    411.43478 

mm_sizes <- table(colData(sfe_tumor)$spatial_community)
mean(mm_sizes)

```



Marker genes for each met 

```{r eval=FALSE}

markers <- data.frame()
for(cur_sample in sample_list){
    print(cur_sample)
    sfe <- sfe_list[[cur_sample]]
    sfe_tumor <- sfe[,sfe$cell_group == 'Tumor']

    # plot vln plots with seurat
    sfe_tumor <- logNormCounts(sfe_tumor)
    X <- as(logcounts(sfe_tumor), 'dgCMatrix')
    seurat_obj <- CreateSeuratObject(
        counts=X, meta.data=as.data.frame(colData(sfe_tumor))
    )
    Idents(seurat_obj) <- seurat_obj$spatial_community

    # loop through each signature:
    cur_markers <- FindAllMarkers(
        seurat_obj,
        slot = 'counts',
        test.use = 'wilcox',
        only.pos=TRUE,
        logfc.threshold = 1.0
    )  %>% subset(p_val_adj <= 0.05)
    cur_markers$sample <- cur_sample 
    markers <- rbind(markers, cur_markers)

}

    head(markers)
    table(markers$cluster)
    length(unique(markers$gene))

markers <- markers[,c('gene', 'cluster', 'sample', 'p_val_adj', 'avg_log2FC')]
colnames(markers)[2] <- 'micromet'

write.table(markers, file='micromet_marker_genes.tsv', sep='\t', row.names=FALSE, quote=FALSE)

```

save the .rds objects for the Zenodo publication

things each .rds needs 
* Micromet ID 
* broad classification (CNS/Imm/Vasc/Tumor)


```{r eval=FALSE}



sample_list <- c('134990_left', '134990_right', '10010900', '10032256', '10058583')
sfe_list <- c()
sfe_list_full <- c()
for(cur_sample in sample_list){
    print(cur_sample)
    x <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
    y <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))

    sfe_list[[cur_sample]] <- x
    sfe_list_full[[cur_sample]] <- y

}

#-----------------------------------------------------------------
# Sample 134990_left 
#-----------------------------------------------------------------

cur_sample <- '134990_left'
sfe_mm <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))

ct_df <- data.frame(
    cluster = c(1:11),
    cell_group = c('Tumor', 'CNS', 'Immune', 'Vasculature', 'CNS', 'Immune', 'CNS', 'Vasculature', 'Vasculature', 'CNS', 'Immune')
)

tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group

# transfer micromet clustering:
tmp1 <- as.data.frame(colData(sfe))
tmp1$bc <- rownames(tmp1)
tmp2 <- as.data.frame(colData(sfe_mm))
tmp2$bc <- rownames(tmp2)

test <- left_join(
    tmp1, 
    tmp2, 
    by = 'bc'
)
colData(sfe)$spatial_community <- test$spatial_community

saveRDS(sfe, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_upload.rds'))



#-----------------------------------------------------------------
# Sample 134990_right
#-----------------------------------------------------------------


cur_sample <- '134990_right'
sfe_mm <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))

ct_df <- data.frame(
    cluster = c(1:13),
    cell_group = c(
        'CNS', 'CNS', 'Tumor', 'Vasculature',
        'CNS', 'Immune', 'Immune', 'Vasculature', 'Immune',
        'CNS', 'Tumor', 'Tumor', 'Vasculature'
        )
)

tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group

# transfer micromet clustering:
tmp1 <- as.data.frame(colData(sfe))
tmp1$bc <- rownames(tmp1)
tmp2 <- as.data.frame(colData(sfe_mm))
tmp2$bc <- rownames(tmp2)

test <- left_join(
    tmp1, 
    tmp2, 
    by = 'bc'
)
colData(sfe)$spatial_community <- test$spatial_community

saveRDS(sfe, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_upload.rds'))


#-----------------------------------------------------------------
# Sample 10010900
#-----------------------------------------------------------------


cur_sample <- '10010900'
sfe_mm <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))

ct_df <- data.frame(
    cluster = c(1:10),
    cell_group = c(
        'Tumor',
        'Immune', # T-cells
        'CNS',
        'Vasculature',
        'Tumor',
        'Tumor',
        'Immune', # T-cells 
        'Immune', # Macrophage (?)
        'Immune', # Mac/Micro (Spi1)
        'Immune' # B cell?
        )
)


tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group

# transfer micromet clustering:
tmp1 <- as.data.frame(colData(sfe))
tmp1$bc <- rownames(tmp1)
tmp2 <- as.data.frame(colData(sfe_mm))
tmp2$bc <- rownames(tmp2)

test <- left_join(
    tmp1, 
    tmp2, 
    by = 'bc'
)
colData(sfe)$spatial_community <- test$spatial_community

saveRDS(sfe, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_upload.rds'))



#-----------------------------------------------------------------
# Sample 10032256
#-----------------------------------------------------------------

cur_sample <- '10032256'
sfe_mm <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))


ct_df <- data.frame(
    cluster = c(1:10),
    cell_group = c(
        'CNS',
        'CNS',
        'Immune', # T cells?
        'CNS', # IDK for sure
        'Vasculature',
        'Immune', # Macrophage
        'Vasculature',
        'Tumor',
        'Tumor', 
        'Immune'
        )
)
tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group

# transfer micromet clustering:
tmp1 <- as.data.frame(colData(sfe))
tmp1$bc <- rownames(tmp1)
tmp2 <- as.data.frame(colData(sfe_mm))
tmp2$bc <- rownames(tmp2)

test <- left_join(
    tmp1, 
    tmp2, 
    by = 'bc'
)
colData(sfe)$spatial_community <- test$spatial_community

saveRDS(sfe, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_upload.rds'))



#-----------------------------------------------------------------
# Sample 10058583
#-----------------------------------------------------------------


cur_sample <- '10058583'
sfe_mm <- readRDS(file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_micromet_clustering.rds'))    
sfe <- readRDS(paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '.rds'))



ct_df <- data.frame(
    cluster = c(1:14),
    cell_group = c(
            'CNS',
            'CNS',
            'CNS',
            'Vasculature',
            'Immune',
            'CNS',
            'Tumor',
            'Immune', # T cells 
            'Tumor',
            'CNS',
            'Immune', # mac
            'CNS',
            'Immune',
            'Immune'
        )
)

tmp <- as.data.frame(colData(sfe))
tmp$cluster <- as.numeric(tmp$cluster)
test <- left_join(
    tmp, 
    ct_df, 
    by = 'cluster'
)
colData(sfe)$cell_group <- test$cell_group

# transfer micromet clustering:
tmp1 <- as.data.frame(colData(sfe))
tmp1$bc <- rownames(tmp1)
tmp2 <- as.data.frame(colData(sfe_mm))
tmp2$bc <- rownames(tmp2)

test <- left_join(
    tmp1, 
    tmp2, 
    by = 'bc'
)
colData(sfe)$spatial_community <- test$spatial_community

saveRDS(sfe, file=paste0(cur_sample, '/data/voyager_sfe_', cur_sample, '_upload.rds'))






```



Meeting notes :

* Calculate the correlation between micromet size (ncells) and mxd4%
all together and with seprate per sample   
    Can I point to specific mets that have differences in size vs mxd4% 
    sample 10058583 is an outlier in the sizes 
    
* correlate signatures as well with the MM size

* Another pie that has any of the 5 markers

* Exclude micromets by size (?)

* pairwise correlation of the micromets + clustering
* within and across patients (!!!)

* strength in numbers -- how many micromets detected ? How many tumor cells? How many cells total?

* for pie charts, make 1 that shows any expressed vs no exprssed, and then make the same as the others 

*

* integrated clustering of these 5 samples