

```{r eval=FALSE}

# interactive 8 256G 10:00:00 genD

library(Seurat)
library(future)
plan("multisession", workers = 8)
library(ggplot2)




setwd('~/collab/Valiente_2024/testing/')
data_dir <- 'data/'
fig_dir <- 'figures/'

# directory for raw 

# get a list of the xenium slides:
raw_dir <- '~/collab/Valiente_2024/data/raw/Xenium_Outs/'
xen_slides <- dir(raw_dir)

# get a list of all samples:
samples <- lapply(xen_slides, function(x){dir(paste0(raw_dir, x))})

cur_sample <- samples[[1]][1]
cur_file <- paste0(raw_dir, '/', xen_slides[1], '/', cur_sample)

# Load the Xenium data
xenium.obj <- LoadXenium(cur_file, fov = "fov")

# remove cells with 0 counts
xenium.obj <- subset(xenium.obj, subset = nCount_Xenium > 0)

p <- ImageDimPlot(
    xenium.obj, 
    fov = "fov", 
    molecules = rownames(xenium.obj)[1:4], 
    nmols = 20000
)

pdf(paste0(fig_dir, 'test_imagedimplot.pdf'), width=5, height=5)
p 
dev.off()




# clustering 
xenium.obj <- NormalizeData(xenium.obj, assay = "Xenium")
xenium.obj <- ScaleData(xenium.obj)
xenium.obj <- RunPCA(xenium.obj, npcs = 30, features = rownames(xenium.obj))
xenium.obj <- RunUMAP(xenium.obj, dims = 1:30)
xenium.obj <- FindNeighbors(xenium.obj, reduction = "pca", dims = 1:30)
xenium.obj <- FindClusters(xenium.obj, resolution = 0.3)


p <- DimPlot(xenium.obj)
pdf(paste0(fig_dir, 'umap_clustering.pdf'), width=10, height=10)
p 
dev.off()

p <- ImageDimPlot(xenium.obj, cols = "polychrome", size = 0.75)

p <- ImageDimPlot(xenium.obj, fov = "fov", axes = TRUE, border.color = "white", border.size = 0.1, cols = "polychrome",
    coord.fixed = FALSE)


pdf(paste0(fig_dir, 'spatial_clustering.pdf'), width=12, height=12)
p 
dev.off()


cropped.coords <- Crop(xenium.obj[["fov"]], x = c(1200, 2900), y = c(3750, 4550), coords = "plot")
xenium.obj[["zoom"]] <- cropped.coords
# visualize cropped area with cell segmentations & selected molecules
DefaultBoundary(xenium.obj[["zoom"]]) <- "segmentation"

p <- ImageDimPlot(xenium.obj, fov = "zoom", axes = TRUE, border.color = "white", border.size = 0.1, cols = "polychrome",
    coord.fixed = FALSE)


pdf(paste0(fig_dir, 'spatial_clustering_zoom.pdf'), width=12, height=12)
p 
dev.off()


```

Try to make a combined object as a test:

Note: 96 GB was not enough memory

Problem with Seurat, taking FOREVER to make a joint object.
There is one outlier sample that has 600k+ cells,
and another with 300k + cells, but many of the samples have 
a few thousand or 10k ish cells. This is making me wonder if it 
makes sense to do a joint analysis at all?

If this is still going by 11:15, I should kill it. I can grab the 
meta-data in the loop and then get the basic QC stats from there. 
Or maybe just start a new tmux instead of kill it.

```{r eval=FALSE}


# interactive 8 256G 10:00:00 genD

library(Seurat)
library(future)
plan("multisession", workers = 8)
library(ggplot2)

# setup directories
setwd('~/collab/Valiente_2024/testing/')
data_dir <- 'data/'
fig_dir <- 'figures/'


# get a list of the xenium slides:
raw_dir <- '~/collab/Valiente_2024/data/raw/Xenium_Outs/'
xen_slides <- dir(raw_dir)

# get a list of all samples:
samples <- lapply(xen_slides, function(x){dir(paste0(raw_dir, x))})

seurat_list <- list() 
i <- 1
for(cur_slide in xen_slides){

    cur_samples <- dir(paste0(raw_dir, cur_slide))
    for(cur_sample in cur_samples){
        print(i)
        print(cur_sample)
        cur_file <- paste0(raw_dir, '/', cur_slide, '/', cur_sample)
        cur_obj <- LoadXenium(cur_file, fov = cur_sample)
        cur_obj$Sample <- cur_sample
        cur_obj$Slide <- cur_slide
        seurat_list[[cur_sample]] <- cur_obj
        i <- i + 1 
    }
}

xenium.obj <- merge(
    seurat_list[[1]], seurat_list[2:length(seurat_list)]
)
dim(xenium.obj)

# save the object!!!
saveRDS(xenium.obj, file=paste0(data_dir, 'xenium_merged_seurat.rds'))


# first sample
cur_sample <- samples[[1]][1]
cur_file <- paste0(raw_dir, '/', xen_slides[1], '/', cur_sample)
xenium.obj1 <- LoadXenium(cur_file, fov = "fov")
xenium.obj1$Sample <- cur_sample

# second sample
cur_sample <- samples[[1]][2]
cur_file <- paste0(raw_dir, '/', xen_slides[1], '/', cur_sample)
xenium.obj2 <- LoadXenium(cur_file, fov = "fov")
xenium.obj2$Sample <- cur_sample
xenium.obj2$Slide <- xen_slides[1]

xenium.obj <- merge(xenium.obj1, xenium.obj2)
dim(xenium.obj)
table(xenium.obj$Sample)

# remove cells with 0 counts
xenium.obj <- subset(xenium.obj, subset = nCount_Xenium > 0)
table(xenium.obj$nCount_Xenium > 0)

p1 <- ImageDimPlot(
    xenium.obj, 
    fov = "fov", 
    molecules = rownames(xenium.obj)[1:4], 
    nmols = 20000
)


p2 <- ImageDimPlot(
    xenium.obj, 
    fov = "fov.2", 
    molecules = rownames(xenium.obj)[1:4], 
    nmols = 20000
)

pdf(paste0(fig_dir, 'test_imagedimplot_merged.pdf'), width=10, height=5)
p1 + p2
dev.off()

```



```{r eval=FALSE}


# interactive 8 256G 10:00:00 genD

library(Seurat)
library(future)
plan("multisession", workers = 8)
library(ggplot2)
library(cowplot)
library(tidyverse)
theme_set(theme_cowplot())

# setup directories
setwd('~/collab/Valiente_2024/testing/')
data_dir <- 'data/'
fig_dir <- 'figures/'


# get a list of the xenium slides:
raw_dir <- '~/collab/Valiente_2024/data/raw/Xenium_Outs/'
xen_slides <- dir(raw_dir)

# get a list of all samples:
samples <- lapply(xen_slides, function(x){dir(paste0(raw_dir, x))})

xen_meta <- data.frame()
i <- 1
for(cur_slide in xen_slides){

    cur_samples <- dir(paste0(raw_dir, cur_slide))
    for(cur_sample in cur_samples){
        print(i)
        print(cur_sample)
        cur_file <- paste0(raw_dir, '/', cur_slide, '/', cur_sample)
        cur_obj <- LoadXenium(cur_file, fov = cur_sample)
        cur_obj$Sample <- cur_sample
        cur_obj$Slide <- cur_slide
        xen_meta <- rbind(xen_meta, cur_obj@meta.data)
        i <- i + 1 
    }
}

# save the object!!!
write.csv(xen_meta, file=paste0(data_dir, 'xenium_combined_meta.csv'), quote=FALSE)

# extract info from the Sample name 
sample_names <- xen_meta$Sample 
tmp <- do.call(rbind, str_split(sample_names, '__'))
xen_meta$xen_slide <- as.factor(tmp[,2])
xen_meta$sample_id <- as.factor(tmp[,3])
xen_meta$exp_date <- tmp[,4] 


# QC Distribution plots:

p <- xen_meta %>% 
    group_by(xen_slide) %>% arrange(sample_id) %>% ungroup %>%
    ggplot(aes(x = sample_id, y = nCount_Xenium, fill=sample_id)) + 
    geom_violin() + 
    geom_boxplot(width=0.1, outlier.shape=NA, fill='white') + 
    RotatedAxis() + 
    NoLegend() + 
    theme(
        panel.grid.major.y=element_line(color='lightgrey', linewidth=0.5)
    ) + xlab('')

pdf(paste0(fig_dir, 'merged_QC_nCountXenium_facet.pdf'), width=8, height=8)
p + facet_wrap(~xen_slide, ncol=2, scales='free')
dev.off()


p <- xen_meta %>% 
    group_by(xen_slide) %>% arrange(sample_id) %>% ungroup %>%
    ggplot(aes(x = sample_id, y = nFeature_Xenium, fill=sample_id)) + 
    geom_violin() + 
    geom_boxplot(width=0.1, outlier.shape=NA, fill='white') + 
    RotatedAxis() + 
    NoLegend() + 
    theme(
        panel.grid.major.y=element_line(color='lightgrey', linewidth=0.5)
    ) + xlab('')

pdf(paste0(fig_dir, 'merged_QC_nFeatureXenium_facet.pdf'), width=8, height=8)
p + facet_wrap(~xen_slide, ncol=2, scales='free')
dev.off()


# number of cells per sample 
plot_df <- xen_meta %>% 
    group_by(xen_slide, sample_id) %>% 
    summarise(count = n()) %>% 
    arrange(-count) 
plot_df$sample_id <- factor(as.character(plot_df$sample_id), levels=as.character(plot_df$sample_id))

p1 <- plot_df %>%
    ggplot(aes(x=sample_id, y = count, fill=sample_id)) + 
    geom_bar(stat='identity') + 
   # facet_wrap(~xen_slide, ncol=2, scales='free') + 
    RotatedAxis() + NoLegend() + xlab('') + ylab('N cells') +
    ggtitle('Number of cells detected per sample') + 
    theme(plot.title=element_text(hjust=0.5))

p2 <- plot_df %>%
    ggplot(aes(x=sample_id, y = count, fill=sample_id)) + 
    geom_bar(stat='identity') + 
    scale_y_log10() + 
    geom_text(aes(y = count, label = count), hjust = 1.5, colour = "black", angle=90) +
   # facet_wrap(~xen_slide, ncol=2, scales='free') + 
    RotatedAxis() + NoLegend() + xlab('') + ylab('N cells') +
    ggtitle('Number of cells detected per sample (log10 scale)') + 
    theme(plot.title=element_text(hjust=0.5))


pdf(paste0(fig_dir, 'merged_QC_n_cells_facet.pdf'), width=8, height=8)
p1 / p2 
dev.off()

p <- xen_meta %>%
    ggplot(aes(x=nCount_Xenium, y=nFeature_Xenium)) +
    geom_bin2d() +
    facet_wrap(~sample_id, ncol=8) + 
    RotatedAxis() + 
    scale_fill_gradient(trans="log")
    
pdf(paste0(fig_dir, 'merged_QC_nCount_vs_nFeature_facet.pdf'), width=12, height=6)
p
dev.off()


```
